---
title: "FA7_Group20"
author: "Group"
date: "2025-04-10"
output:
  pdf_document: default
  html_document: default
---

[Click to go to Github link](https://github.com/pywen1004/FA7_GROUP20.git)

Chosen Scenario:

Our group selected the frequency of students tapping their feu identification cards upon entering the Far Eastern University (FEU) electronics library as a practical campus-related issue. In this scenario, the events of students tapping in are random and occur over time. By applying the exponential distribution, we aim to analyze the waiting time between each tap-in event.

Data Collection Process:

For our data collection, we conducted a 30-minute observation period at the counter of the FEU electronics library. During this time, we recorded the number of students who tapped their student IDs to enter the E-library. Each time a student tapped in, we started a stopwatch to measure the time interval between successive taps. The data collected represents the waiting times between student entries, which will be analyzed using the exponential distribution. This process will help us understand how frequently students enter the E-library and the average waiting time between their entries.

```{r, echo=FALSE}
intervals <- c(
  "00:01:24.359", "00:00:27.575", "00:00:44.872", "00:01:14.810", "00:00:02.989",
  "00:00:06.521", "00:00:02.769", "00:01:11.726", "00:02:06.428", "00:00:28.773",
  "00:00:53.327", "00:00:12.752", "00:00:07.608", "00:00:02.713", "00:01:06.887",
  "00:00:04.112", "00:01:20.369", "00:00:03.009", "00:00:07.110", "00:02:06.023",
  "00:00:06.136", "00:00:15.328", "00:00:26.034", "00:00:04.230", "00:00:57.618",
  "00:00:04.030", "00:00:10.538", "00:00:31.063", "00:00:05.023", "00:00:55.850",
  "00:01:15.879", "00:01:47.911", "00:00:36.460", "00:00:09.203", "00:00:06.780",
  "00:00:04.845", "00:01:07.385", "00:01:26.927", "00:00:13.199", "00:00:50.096",
  "00:01:12.152", "00:01:45.032", "00:00:49.731"
)

time_to_minutes <- function(x) {
  parts <- as.numeric(unlist(strsplit(x, ":")))
  return((parts[1]*3600 + parts[2]*60 + parts[3]) / 60)
}
interval_minutes <- sapply(intervals, time_to_minutes)

mean_time <- mean(interval_minutes)
lambda <- 1 / mean_time
```

#### Average time in minutes:

```{r, echo=FALSE}
cat("Mean time (minutes):", mean_time, "\n")
```

#### Rate per minute:

```{r, echo=FALSE}
cat("Lambda (rate per minute):", lambda, "\n")
```

#### Student taps per minute:

```{r, echo=FALSE}
prob_less_than_1 <- pexp(1, rate = lambda)
cat("P(Student taps in within 1 minute):", prob_less_than_1, "\n")
```

```{r, echo=FALSE}
hist(interval_minutes, probability = TRUE, main = "Time Between Tap-Ins (FEU Library)",
     xlab = "Minutes", col = "skyblue", breaks = 10)
curve(dexp(x, rate = lambda), col = "red", lwd = 2, add = TRUE)
```

#### PDF

```{r, echo=FALSE}
x <- 1
pdf_value <- dexp(x, rate = lambda)
cat("PDF at x = 1 minute:", pdf_value, "\n")

curve(dexp(x, rate = lambda), from = 0, to = 5,
      main = "Exponential PDF (Tap-in Wait Times)",
      xlab = "Minutes", ylab = "Density",
      col = "red", lwd = 2)
```

#### CDF

```{r, echo=FALSE}
cdf_1min <- pexp(1, rate = lambda)
cat("CDF at x = 1 minute:", cdf_1min, "\n")
curve(pexp(x, rate = lambda), from = 0, to = 5,
      main = "Exponential CDF (Tap-in Wait Times)",
      xlab = "Minutes", ylab = "Cumulative Probability",
      col = "blue", lwd = 2)
```

#### Q-Q Plot

```{r, echo=FALSE}
qqplot(qexp(ppoints(length(interval_minutes)), rate = lambda),
       interval_minutes,
       main = "Q-Q Plot: Observed vs Exponential",
       xlab = "Theoretical Quantiles", ylab = "Observed Quantiles")
abline(0, 1, col = "red", lwd = 2)
```

#### Statistical Goodness-of-Fit Test

```{r, echo=FALSE}
ks.test(interval_minutes, "pexp", rate = lambda)
```

Interpretations and real-world implications:

With the data given we can see that in the electronics library numerous students tapped in within a very small amount of time between each other. Thus implying that the electronics library is a place that very easily and very quickly fills up within a short amount of time. Due to the mean interval being about half a minute long and not instantaneous this means that long lines aren’t necessarily an issue, but the frequency to which it happens means that a full electronics library will become an issue. This data can help both librarian staff and students better prepare and plan ahead when visiting the electronics library especially during peak hours. This can also help the University Administrators gauge the interest of students in the electronics library, possibly leading to bigger/improved spaces, longer library hours, and etc.
